---
title: "Regular Expressions in R"
author: "Andrew Yeo"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("C:/Users/Andrew/Desktop/Coding/R")
library(stringr)
```

We can use some string functions to extract filenames; documents on dplyr. 
grep(): If parameter value = TRUE, it returns the matches whereas value = FALSE returns their indices. The invert argument let’s you get everything BUT the pattern you specify. 
grepl(): returns a logical vector. See here for more information.

```{r}
files=list.files()
str(files)
grep("Dplyr", files, value = TRUE)
grep("Dplyr", files, value = TRUE, invert = TRUE)
grep("Dplyr", files, value = FALSE)
grep("Dplyr", files, value = FALSE, invert = TRUE)

grepl("Dplyr", files)
```


Regular Expression (Regex) is a pattern that describes a specific set of strings with a common structure. It is heavily used for string matching / replacing in all programming languages, although specific syntax may differ a bit. 

* identify match to a pattern: grep(..., value = FALSE), grepl(), stringr::str_detect()
* extract match to a pattern: grep(..., value = TRUE), stringr::str_extract(),  stringr::str_extract_all()
* locate pattern within a string, i.e. give the start position of matched patterns. regexpr(), gregexpr(),  stringr::str_locate(), string::str_locate_all()
* replace a pattern: sub(), gsub(), stringr::str_replace(), stringr::str_replace_all()
* split a string using a pattern: strsplit(), stringr::str_split()

Regular expressions typically specify characters to seek out; information about repeats within the string. This is accomplished with the help of metacharacters that have specific meaning: $ * + . ? [ ] ^ { } | ( ) 


## Escape 
Sequences. List  of things that cannot be directly coded in a string:

* \': single quote. You don’t need to escape single quote inside a double-quoted string, so we can also use "'" in the previous example.
* \": double quote. Similarly, double quotes can be used inside a single-quoted string, i.e. '"'.
* \n: newline.
* \r: carriage return.
* \t: tab character.

```{r}
gDat <- read.delim("gapminderDataFiveYear.txt")
head(gDat)
grep('\'', levels(gDat$country), value = TRUE)  # note how we had to to use \ to escape 
grep("'", levels(gDat$country), value = TRUE)

print("a\nb")
paste("hello", "wor\nld")
cat("a\nb")                # if printing a string using these sequences, use cat().
```

## Quantifiers
When we want to match a certain number of characters that meet a certain criteria we can apply quantifiers to our pattern searches

* *: matches at least 0 times.
* +: matches at least 1 times.
* ?: matches at most 1 times.
* {n}: matches exactly n times.
* {n,}: matches at least n times.
* {n,m}: matches between n and m times.

```{r}
strings <- c("a", "ab", "acb", "accb", "acccb", "accccb")

grep("ac*b", strings, value = TRUE)     #The c has to match 0 or more times
grep("ac+b", strings, value = TRUE)     #The c has to match 1 or more times
grep("ac?b", strings, value = TRUE)
grep("ac{2}b", strings, value = TRUE)
grep("ac{2,}b", strings, value = TRUE)  #The c has to match >2 times
grep("ac{2,3}b", strings, value = TRUE) 

grep("ee", gDat$country,value=TRUE )
unique(grep("ee", gDat$country,value=TRUE ))
unique(grep("ee", gDat$country,value=TRUE, invert = TRUE )) # everything but Greek

```

## Position of pattern within a string

* ^: matches the start of the string.
* $: matches the end of the string.
* \b: matches the empty string at either edge of a word. Don’t confuse it with ^ $ which marks the edge of a string.
* \B: matches the empty string provided it is not at an edge of a word.

```{r}
(strings <- c("abcd", "cdab", "cabd", "c abd"))
grep("ab", strings, value = TRUE)
grep("^ab", strings, value = TRUE)
grep("ab$", strings, value = TRUE)
grep("\\bab", strings, value = TRUE)

files=list.files()
grep(".txt$",files, value=TRUE)
```

## Operators

* .: matches any single character, as shown in the first example.
* [...]: a character list, matches any one of the characters inside the square brackets. We can also use  - inside the brackets to specify a range of characters.
* [^...]: an inverted character list, similar to [...], but matches any characters except those inside the square brackets.
* \: suppress the special meaning of metacharacters in regular expression, i.e.  $ * + . ? [ ] ^ { } | ( ) \, similar to its usage in escape sequences. Since \ itself needs to be escaped in R, we need to escape these metacharacters with double backslash like \\$.
* |: an “or” operator, matches patterns on either side of the |.
* (...): grouping in regular expressions. This allows you to retrieve the bits that matched various parts of your regular expression so you can alter them or use them for building up a new string. Each group can than be refer using \\N, with N being the No. of (...) used. This is called backreference.

```{r}
(strings <- c("^ab", "ab", "abc", "abd", "abe", "ab 12"))
grep("ab.", strings, value = TRUE)
grep("ab[c-e]", strings, value = TRUE)
grep("ab[^c]", strings, value = TRUE)
grep("^ab", strings, value = TRUE)
grep("\\^ab", strings, value = TRUE)
grep("abc|abd", strings, value = TRUE)
gsub("(ab) 12", "\\1 34", strings)
```

## Character Classes
Character classes allows to – surprise! – specify entire classes of characters, such as numbers, letters, etc. There are two flavors of character classes, one uses [: and :] around a predefined name inside square brackets and the other uses \ and a special character. They are sometimes interchangeable:

* [:digit:] or \d: digits, 0 1 2 3 4 5 6 7 8 9, equivalent to [0-9].
* \D: non-digits, equivalent to [^0-9].
* [:lower:]: lower-case letters, equivalent to [a-z].
* [:upper:]: upper-case letters, equivalent to [A-Z].
* [:alpha:]: alphabetic characters, equivalent to [[:lower:][:upper:]] or [A-z].
* [:alnum:]: alphanumeric characters, equivalent to [[:alpha:][:digit:]] or [A-z0-9].
* \w: word characters, equivalent to [[:alnum:]_] or [A-z0-9_].
* \W: not word, equivalent to [^A-z0-9_].
* [:xdigit:]: hexadecimal digits (base 16), 0 1 2 3 4 5 6 7 8 9 A B C D E F a b c d e f, equivalent to  [0-9A-Fa-f].
* [:blank:]: blank characters, i.e. space and tab.
* [:space:]: space characters: tab, newline, vertical tab, form feed, carriage return, space.
* \s: space, ` `.
* \S: not space.
* [:punct:]: punctuation characters, ! " # $ % & ’ ( ) * + , - . / : ; < = > ? @ [  ] ^ _ ` { | } ~.
* [:graph:]: graphical (human readable) characters: equivalent to [[:alnum:][:punct:]].
* [:print:]: printable characters, equivalent to [[:alnum:][:punct:]\\s].
* [:cntrl:]: control characters, like \n or \r, [\x00-\x1F\x7F].  
Note:
* [:...:] has to be used inside square brackets, e.g. [[:digit:]].
* \ itself is a special character that needs escape, e.g. \\d. Do not confuse these regular expressions with R escape sequences such as \t.


```{r}

```

