---
title: "R Markdown- Intro"
output: html_document
---
## Intro 
Render:Process of knitting and converting
knit through "Knitr": knitr runa each chunk of R code in the document and append the results of the code to the document next to the code chunk. Each report contains the code it needs to make its own graphs, tables, numbers, etc. The author can automatically update the report by re-knitting.

-convert: The rmarkdown package uses the pandoc program to transform the file into a new format. Converting your .Rmd file into an HTML, PDF, or Microsoft Word file. 

 R Markdown reports rely on three frameworks:

* markdown for formatted text
* knitr for embedded R code
* YAML for render parameters


# Basics
### Markdown for formatted text
You can use markdown to indicate

* **bold** and *italic*: text have to be without spaces
* lists:  Group lines into bullet points that begin with asterisks. Leave a blank line before the first bullet
* headers (e.g., section titles) are created with # or ##
* hyperlinks: [Github](www.github.com) or [StackOverflow](www.stackoverflow.com)

### Rendering
To transform your markdown file into an HTML, PDF, or Word document, click the “Knit” icon that appears above your file in the scripts editor.



# Knitr
The knitr package extends the basic markdown syntax to include chunks of executable R code.

### Embedded R code
To embed a chunk of R code into your report, surround the code with two lines that each contain three backticks. After the first set of backticks, include {r}, which alerts knitr that you have included a chunk of R code.

* include = FALSE prevents code and results from appearing in the finished file. R Markdown still runs the code in the chunk, and the results can be used by other chunks.
* echo = FALSE prevents code, but not the results from appearing in the finished file. This is a useful way to embed figures.
* message = FALSE prevents messages that are generated by code from appearing in the finished file.
* warning = FALSE prevents warnings that are generated by code from appearing in the finished.
* fig.cap = "..." adds a caption to graphical results.


```{r }
# Displays code and displays results
dim(iris)
```
```{r eval=FALSE}
# Displays code and omits results
dim(iris)
```
```{r echo=FALSE}
# Omits code and displays results
# useful for plots
dim(iris)
```

### Inline code
To embed R code in a line of text, surround the code with a pair of backticks and the letter r, like this:
For example, the following is a inline code: 4 times 5 equals `r 4*5`. Note that R markdown always displays the results of inline code but not the code. 



# YAML for render parameters
You can use a YAML header to control how rmarkdown renders your .Rmd file. A YAML header is a section of key: value pairs surrounded by --- marks as shown at the top of the document.

* title 
* author
* date
* output- must be "html_document", "pdf_document", "word_document"

Parameters are declared using the params field within the YAML header of the document. 