---
title: "DateTime Objects"
output: html_notebook
---

# Datetime Objects
Internally, Date objects are stored as the number of days since January 1, 1970, using negative numbers for earlier dates.  
* The as.numeric function can be used to convert a Date object to its internal form.
* To extract the components of the dates, the weekdays, months, days or quarters functions can be used. 

Code	Value: 
%d	Day of the month (decimal number)

%m	Month (decimal number)
%b	Month (abbreviated)
%B	Month (full name)

%y	Year (2 digit)
%Y	Year (4 digit)

```{r Basic DateTime Creation}
as.Date("2012-08-30")
as.Date("30/08/2012", format = "%d/%m/%Y")
as.Date('April 26, 2001',format='%B %d, %Y')
as.Date('22JUN01', format='%d%b%y')
as.Date('1/15/2001',format='%m/%d/%Y')

class(as.Date('1/15/2001',format='%m/%d/%Y'))
Sys.Date() - as.Date("1970-01-01")
difftime(Sys.Date(), as.Date("1970-01-01"), units = "days")
difftime(Sys.Date(), as.Date("1970-01-01"), units = "weeks")

```


```{r Datetime lists}
bdays = c(tukey=as.Date('1915-06-16'), fisher=as.Date('1890-02-17'),
          cramer=as.Date('1893-09-25'), kendall=as.Date('1907-09-06'))
          
weekdays(bdays)
months(bdays)
quarters(bdays)
```

```{r Current Date and Time}
date=Sys.Date()
time=Sys.time()

date
date+1  #add an additional day

time
time+1        #add an aditional second
time+60       #minute
time+60*60    #hour
time+60*60*24 #day
```


# Lubridate Objects ----
To format dates, lubridate provides a series of functions that are a permutation of the letters “m”, “d” and “y” to represent the ordering of month, day and year. For example, if our data has a column of dates such as May 11, 1996, our dates are ordered month-day-year. Therefore we would use the mdy function to transform the column to a date object. If our dates were in the order of, say, year-month-day, we would use the ymd function. lubridate provides functions for every permutation of “m”, “d”, “y”.

```{r}
library(lubridate)
# Note how the dates are immediately converted into date classes
begin <- c("May 11, 1996", "September 12, 2001", "July 1, 1988")
class(begin)
begin <- mdy(begin)   ## [1] "1996-05-11" "2001-09-12" "1988-07-01"
class(begin)

ymd("20110604")
ymd("2011/06/04")
dmy("04/06/2011")


mdy_hm("3-11-2017 5:21", tz = "US/Eastern")
ymd_hms('2011-04-06 08:00:00')
ymd_hms("2011-06-04 12:00:00", tz = "Pacific/Auckland")
ymd_hms("2011-08-10 14:00:00", tz = "Pacific/Auckland")

# Dealing with time
time1 <- c("1:13", "0:58", "1:01")
time1 <- ms(time1) ## [1] "1M 13S" "58S"    "1M 1S"
time2 <- c("12:23:11", "09:45:31", "12:05:22")
time2 <- hms(time2) ## [1] "12H 23M 11S" "9H 45M 31S"  "12H 5M 22S"
time3 <- c("2:14", "2:16", "3:35")
time3 <- hm(time3) ## [1] "2H 14M 0S" "2H 16M 0S" "3H 35M 0S"

```

Rounding Dates to the closest number:  
```{r}
x <- ymd_hms("2009-08-03 12:01:59.23")
round_date(x, ".5s")
round_date(x, "sec")
round_date(x, "second")
round_date(x, "minute")
round_date(x, "5 mins")
round_date(x, "hour")
round_date(x, "2 hours")
round_date(x, "day")
round_date(x, "week")
round_date(x, "month")
round_date(x, "bimonth")
round_date(x, "quarter") == round_date(x, "3 months")
round_date(x, "halfyear")
round_date(x, "year")
```

Using Ceiling dates and floor dates. Ceiling dates and floor dates is the beginning and the end of the specified interval. 
```{r}
x <- ymd_hms("2009-08-03 12:01:59.23")
floor_date(x, ".1s")
floor_date(x, "second")
floor_date(x, "minute")
floor_date(x, "hour")
floor_date(x, "day")
floor_date(x, "week")
floor_date(x, "month")
floor_date(x, "bimonth")
floor_date(x, "quarter")
floor_date(x, "season")
floor_date(x, "halfyear")
floor_date(x, "year")

ceiling_date(x, ".1 sec") # imprecise representation at 0.1 sec !!!
ceiling_date(x, "second")
ceiling_date(x, "minute")
ceiling_date(x, "5 mins")
ceiling_date(x, "hour")
ceiling_date(x, "day")
ceiling_date(x, "week")
ceiling_date(x, "month")
ceiling_date(x, "bimonth") == ceiling_date(x, "2 months")
ceiling_date(x, "quarter")
ceiling_date(x, "season")
ceiling_date(x, "halfyear")
ceiling_date(x, "year")
```

Modifying dates:
```{r Modifying POSIX}
a <- ymd_hms("2009-08-03 12:01:59.23"); month(a)
month(a) = 12; a

second(a)
second(a) = 12; a
second(a) = 400 # result rolls over into the next higher element:

update(a, day = 2, month = 3, second = 24) #update multiple values
```

Time Intervals:
Since we’re dealing with elapsed time between two dates, let’s start with Intervals. We can define an Interval using the %--% operator:
* They show the beginng date and end date
```{r}
start <- mdy_hm("3-11-2017 5:21", tz = "US/Eastern")
end <- mdy_hm("3-12-2017 5:21", tz = "US/Eastern")
time.interval <- start %--% end
str(time.interval)  #  notice how the time zone changes from EST to EDT; indicates  Daylight Savings started
```

TIme Duration:
To create a Duration between these two dates, we can use the as.duration function. Because Daylight Savings went into effect at 2:00 AM during the interval, an hour was skipped. Thus the duration between these two time points is only 23 hours.
```{r}
time.duration <- as.duration(time.interval)
time.duration
str(time.duration)
time.period <- as.period(time.interval)

```

Calculations and conversions
If we view the elapsed.time object we’ll just see date ranges. We can use as.duration or even as.numeric to view the elapsed time in seconds, weeks or days. 

```{r}
start <- c("2012-08-21", "2012-09-01", "2012-08-15", "2012-09-18")
end <- c("2012-09-16", "2012-09-06", "2012-08-22", "2012-10-11")
elapsed.time <- start %--% end

as.duration(elapsed.time) / dweeks(1)
as.duration(elapsed.time) / dhours(1)
as.duration(elapsed.time) / ddays(1)
as.duration(elapsed.time) / dminutes(1)
as.duration(elapsed.time) / dyears(1)

mean(as.duration(elapsed.time) / dweeks(1))

```

