---
title: "R Notebook"
output: html_notebook
---




# UI ----
Input Functions; Any sort of interaction a user can make is an input
Our slider input. When we run, we see a the html code. 
Input functions take many arguments: input name, label,... ...

```{r}
sliderInput(inputId="num",              # Argument Name
            label="Choose a number",    # The explanation of an input. We can also do "" 
            value=25, min=1, max=100)   # our different input specific arguments
```

Output Functions; An output that places into the app different outputs
```{r}
plotOutput(outputId="hist")   # unique name of an output
```

# Server function ----
Helps assemble input into outputs.
1) Save the output object with a dollar sign: output$hist
2) render function creates the type of output we wish to make. 
```{r}
renderPlot({hist(rnorm(100)) })
```
3) Use input values with input$; input$name shows the input that users choose
```{r}
server= function(input, output){
  output$hist = renderPlot({
    title= "100  random normal values"
    hist(rnorm(100), main=title)
  })
}

server= function(input, output){
  output$hist = renderPlot({
    hist(rnorm(input$num))
  })
}
```

# Sharing ShinyApps
Every webserver needs an app.R which will build the app. 
Also ensure to include all the different datasets, images, ss, helper scripts in the same folder. 
We can also save it in 2 file apps; ui.R and server.R. So the file names will be server.R and ui.R
app.R -> server.R & ui.R


Go to shinyapps.io and we can publish our own shiny apps. 
Install a package called shinyapps through github:
- In r enviroment, do: devtools:install_github("rstudio/shinyapps")
- After installing the app, create an account and get the token from the page.
- Everytime, we can publish the app onto the account with the button on top


We can also build our own server! 