---
title: "R Notebook"
---
# Basic Line Plots
Create line plots with points. Note that they the data points have to be in the same group to be "pathed"
```{r}
library(ggplot2)
df <- data.frame(dose=c("D0.5", "D1", "D2"),len=c(4.2, 10, 29.5)); head(df)

# Basic line plot with points
ggplot(data=df, aes(x=dose, y=len))+geom_line()+geom_point()
ggplot(data=df, aes(x=dose, y=len, group=1))+geom_line()+geom_point()
ggplot(data=df, aes(x=dose, y=len, group=1))+geom_line(linetype = "dashed")+ geom_point()   # Change the line type
ggplot(data=df, aes(x=dose, y=len, group=1))+geom_line(color="red")+ geom_point()  # Change the color

# Adding arrows 
ggplot(data=df, aes(x=dose, y=len, group=1))+geom_line(arrow = arrow())+geom_point()

myarrow=arrow(angle = 30, ends = "both", type = "closed")   # Add a closed arrow to the end of the line
# angle corresponds to the size of the arrow
ggplot(data=df, aes(x=dose, y=len, group=1)) +
  geom_line(arrow=myarrow)+ geom_point()
```

Observations can be also connected using the functions geom_step() or geom_path() :
* geom_line : Connecting observations, ordered by x value
* geom_path() : Observations are connected in original order
* geom_step : Connecting observations by stairs

```{r}
ggplot(data=df, aes(x=dose, y=len, group=1)) +geom_step()+geom_point()
ggplot(data=df, aes(x=dose, y=len, group=1)) +geom_path()+geom_point()
```


# Line plot with multiple groups
In the graphs below, line types and point shapes are controlled automatically by the levels of the variable supp. It is also possible to change manually the line types using the function scale_linetype_manual().

```{r}
df2 <- data.frame(supp=rep(c("VC", "OJ"), each=3),dose=rep(c("D0.5", "D1", "D2"),2),
                len=c(6.8, 15, 33, 4.2, 10, 29.5)); head(df2)

# Create line plots
ggplot(data=df2, aes(x=dose, y=len, group=supp)) + geom_line()+ geom_point()     # Line plot with multiple groups

ggplot(data=df2, aes(x=dose, y=len, group=supp)) +                               # Change line types
  geom_line(linetype="dashed", color="blue", size=1.2)+  geom_point(color="red", size=3)

# Change line types by groups
ggplot(df2, aes(x=dose, y=len, group=supp)) +geom_line(aes(linetype=supp))+geom_point()

# Change line types and point shapes
ggplot(df2, aes(x=dose, y=len, group=supp)) +geom_line(aes(linetype=supp))+ geom_point(aes(shape=supp))

# Changing line types manually
ggplot(df2, aes(x=dose, y=len, group=supp)) +geom_line(aes(linetype=supp))+ geom_point()+
  scale_linetype_manual(values=c("twodash", "dotted"))

```

## Change line colors by graphs
Line colors are controlled automatically by the levels of the variable supp.

It is also possible to change manually line colors using the functions :
* scale_color_manual() : to use custom colors
* scale_color_brewer() : to use color palettes from RColorBrewer package
* scale_color_grey() : to use grey color palettes
```{r}
p<-ggplot(df2, aes(x=dose, y=len, group=supp)) +
  geom_line(aes(color=supp))+
  geom_point(aes(color=supp)); p

# Use custom color palettes
p+scale_color_manual(values=c("#999999", "#E69F00", "#56B4E9"))
# Use brewer color palettes
p+scale_color_brewer(palette="Dark2")
# Use grey scale
p+scale_color_grey() + theme_classic()

```


## Line Plot with a numeric or dates on the x-axis
If the variable on x-axis is numeric, it can be useful to treat it as a continuous or a factor variable depending on what you want to do :
```{r}
# Create some data
df2 <- data.frame(supp=rep(c("VC", "OJ"), each=3),
                dose=rep(c("0.5", "1", "2"),2),
                len=c(6.8, 15, 33, 4.2, 10, 29.5))
head(df2)

# x axis treated as continuous variable
df2$dose <- as.numeric(as.vector(df2$dose))
ggplot(data=df2, aes(x=dose, y=len, group=supp, color=supp)) +
  geom_line() + geom_point()+
  scale_color_brewer(palette="Paired")+
  theme_minimal()
# Axis treated as discrete variable
df2$dose<-as.factor(df2$dose)
ggplot(data=df2, aes(x=dose, y=len, group=supp, color=supp)) +
  geom_line() + geom_point()+
  scale_color_brewer(palette="Paired")+
  theme_minimal()
```





