# Admin ----
```{r}
library(reshape2)
library(plyr)

head(airquality)
str(airquality)
```

Usage and Arguments of melt function
- data: data frame to melt
- id.vars: This is something like our group by variable; using the vector of id variables. Can be integer (variable position) or string (variable name). If blank, will use all non-measured variables.
- measure.vars: vector of measured variables. Can be integer (variable position) or string (variable name). If blank, will use all non id.vars

Naming Convention variables: 
- variable.name	name of variable used to store measured variable names
- value.name	name of variable used to store values

Specifications: 
- na.rm	Should NA values be removed from the data set? This will convert explicit missings to implicit missings.
- factorsAsStrings	Control whether factors are converted to character when melted as measure variables. When FALSE, coercion is forced if levels are not identical across the measure.vars.


## Wide to Long Format
```{r }
# Melt 1= melt assumes all columns with numeric values are variables with values
aql <- melt(airquality)                                        
head(aql)



# Melt 2= Grouping by Month and Day, want to know the values of ozone, solar.r, wind, and temp
aql2_1 <- melt(airquality, id.vars = c("Month", "Day"))         
head(aql2_1)

aql2_2 <- melt(airquality, id.vars = c("Month", "Day"),
             variable.name = "climate_variable", value.name = "climate_value") #naming our variables and values
head(aql2_2)



# Melt 3= Filtering to show only ozone and temp as our variables
aql3_1 <- melt(airquality, id.vars = c("Month", "Day"), measure.vars=c("ozone",'temp'),  
  variable.name = "climate_variable", value.name = "climate_value")
head(aql3_1)

```

## Long to Wide Format
dcast uses a formula to describe the shape of the data. The arguments on the left refer to the ID variables and the arguments on the right refer to the measured variables.

```{r}
# Melt+Dcast 1= 
aql4 <- melt(airquality, id.vars = c("Month", "Day"))
aqw4 <- dcast(data=aql4, Month + Day ~ variable)
head(aqw4)

```

More things we could do with dcast: 
```{r}
aql4 <- melt(airquality, id.vars = c("Month", "Day"))
aqw5=dcast(data=aql4, Month + Day ~ variable, 
           subset = .(variable == 'temp')) #library plyr needed
head(aqw5)

aqw6=dcast(data=aql4, Month ~ variable, 
           fun.aggregate = max, na.rm = TRUE, margins = T)
head(aqw6)

aqw7=dcast(data=aql4, Month ~ variable, 
           fun.aggregate = mean, na.rm = TRUE, margins = T)
head(aqw7)
```


### Aggregation
. is used when there is more than one value per data cell. For example,
```{r}
str(airquality)
aql <- melt(airquality, id.vars = c("Month", "Day"))
str(aql)
dcast(aql, Month ~ variable)
dcast(aql, Month ~ variable, fun.aggregate = mean, na.rm = TRUE)
```

