---
title: "ARCH and GARCH Modelling"
---

# Admin
The walkthrough can be found at: [ARCH&GARCH Tute](https://www.youtube.com/watch?v=8VXmRl5gzEU&t=797s)
```{r}
library(quantmod)
library(rugarch)
library(rmgarch)
```

# Data Upload 
"quantmod" package: A data retrieval function (getSymbols)in order to retrieve some data. This function works, for instance, to retrieve stock data; default source is Yahoo Finance. 

You can see that this object contains a range of daily observations (Open, High, Close, Volume and Adjusted share price). We also learn that the object is formatted as an xts object. 
```{r}
#Specify period of time we are interested in
startDate = as.Date("2007-01-03"); endDate = as.Date("2018-04-30")
 
getSymbols("IBM", from = startDate, to = endDate)
getSymbols("GOOG", from = startDate, to = endDate)
getSymbols("BP", from = startDate, to = endDate)

head(IBM)
str(IBM)
chartSeries(GOOG)

```

When we are estimating volatility models we work with returns. There is a function that transforms the data to returns. There is also a weeklyReturn function! 
 
```{r}
rIBM <- dailyReturn(IBM)
rBP <- dailyReturn(BP)
rGOOG <- dailyReturn(GOOG)

# We put all data into a data frame for use in the multivariate model
rX <- data.frame(rIBM, rBP, rGOOG)
names(rX)[1] <- "rIBM"
names(rX)[2] <- "rBP"
names(rX)[3] <- "rGOOG"
```

# Univariate GARCH Model
Here we are using the functionality provided by the "rugarch" package written by Alexios Galanos.

## Model Specification
The first thing you need to do is to ensure you know what type of GARCH model you want to estimate and then let R know about this: the ugarchspec( ) function which is used to let R know about the model type.

The key issues here are the spec for the Mean Model (here an ARMA(1,1) model) and the specification for the GARCH Model, here an sGARCH(1,1) which is basically a GARCH(1,1). To get details on all the possible specifications and how to change them it is best to consult the documentation of the rugarch package.

```{r}
# Default is an ARMA(1,1)
ug_spec = ugarchspec(); ug_spec
# Say want to change the mean model from an ARMA(1,1) to an ARMA(1,0)
ug_spec <- ugarchspec(mean.model=list(armaOrder=c(1,0))); ug_spec

#more complicated example: EWMA Model
ewma_spec = ugarchspec(variance.model=list(model="iGARCH", garchOrder=c(1,1)),
                       mean.model=list(armaOrder=c(0,0), include.mean=TRUE),  
                       distribution.model="norm", fixed.pars=list(omega=0))
```
## Model Estimation
Now that we have specified a model to estimate we need to find the best parameters, i.e. we need to estimate the model. This step is achieved by the ugarchfit function. 

```{r}
ugfit = ugarchfit(spec = ug_spec, data = rIBM)
ugfit # ugfit is now a list that contains a range of results from the estimation. 
```

The object ugfit contains all the statistical information. In that object you can find two drawers; @fit and @model. Each of these drawers contains a range of different things. 

```{r}
names(ugfit@model)
names(ugfit@fit)

ugfit@fit$coef
ug_var <- ugfit@fit$var              # save the estimated conditional variances
ug_res2 <- (ugfit@fit$residuals)^2   # save the estimated squared residuals

plot(ug_res2, type = "l"); lines(ug_var, col = "green")
```


## Model Forecasting
The function used to subsequently forecast the conditional variance. The function used for this purpose is the "ugarchforecast" function from the rugarch package. 
As you can see we have produced forecasts for the next ten days, both for the expected returns (Series) and for the conditional volatility. Similar to the object created for model fitting, ugfore contains two slots (@model and @forecast) and you can use names(ugfore@forecast) to figure out under which names the elements are saved. 

```{r}
ugfore <- ugarchforecast(ugfit, n.ahead = 10); ugfore

# For instance you can extract the conditional volatility forecast as follows. 
# FOr 10 periods ahead, there is decreasing volatility
ug_f <- ugfore@forecast$sigmaFor
plot(ug_f, type = "l")
```

Note that the volatility is the square root of the conditional variance. To put these forecasts into context let's display them together with the last 50 observations used in the estimation.
You can see how the forecast of the conditional variance picks up from the last estimated conditional variance. In fact it decreases from there, slowly, towards the unconditional variance value.

```{r}
ug_var_t <- c(tail(ug_var,20),rep(NA,10))  # gets the last 20 observations
ug_res2_t <- c(tail(ug_res2,20),rep(NA,10))  # gets the last 20 observations
ug_f <- c(rep(NA,20),(ug_f)^2)

plot(ug_res2_t, type = "l")
lines(ug_f, col = "orange")
lines(ug_var_t, col = "green")
```

