---
title: "Dplyr Package 101_1"
---
From: https://cran.r-project.org/web/packages/dplyr/vignettes/dplyr.html
Data: nycflights13

Basics
- filter() to select cases based on their values.
- arrange() to reorder the cases.
- select() and rename() to select variables based on their names.
- mutate() and transmute() to add new variables that are functions of existing variables.
- summarise() to condense multiple values to a single value.

```{r Admin}
library(dplyr); library(nycflights13)
dim(flights); flights
```

# Filter, Selecting and Mutating
```{r Filter and Arrange}
flights[flights$month == 1 & flights$day == 1, ]  # Base R Code

filter(flights, month == 1) # Using Dplyr Instead: 
filter(flights, month == 1 & day == 1)

arrange(flights, year)
arrange(flights, year, month, day)
arrange(flights, desc(arr_delay))
```

Things we learned: 
starts_with or ends_with to filter column name with a certain name
```{r Select_1}
colnames(flights)
select(flights, everything())

select(flights, year)
select(flights, year:day)              # Select all columns between year and day (inclusive)
select(flights, year, month, day)      # Select columns by name
select(flights, year, month, arr_time:arr_delay)

select(flights, -(year:day))           # Select all columns except those from year to day (inclusive)
select(flights, tail_num = tailnum) 

select(flights, -starts_with("d"), -ends_with("time")) # can't start with "d" or ends with "time"
```

Things we learned:
- identity: In a bracket, Dplyr reads column names as variable instead of string. Identity helps pull our the value instead of using the term as a column name. 
```{r Select_2}
# Part 1
colnames(flights)
select(flights, year)
select(flights, 1)  # select column using the column index
year <- "dep"
select(flights, starts_with(year))  # columns that start with "dep"

# Part 2
select(flights, "year", "month", "day")
year <- 5
select(flights, year, identity(year))  # identity helps pull out the value 
select(flights, year, !! year)         # alternatively, use the !!
```


Things we learned: 
transmute: If you only want to keep the new variables, use transmute():
```{r Mutate}
mutate(flights, gain = arr_delay - dep_delay, speed = distance / air_time * 60)

transmute(flights, gain = arr_delay - dep_delay,
  gain_per_hour = gain / (air_time / 60)) 
```

# Group by and Summarising
```{r Summarizing}
daily <- group_by(flights, year, month, day)

per_day   <- summarise(daily, flights = n())
per_month <- summarise(per_day, flights = sum(flights))
per_year  <- summarise(per_month, flights = sum(flights))
```

```{r GroupBy_1}
by_tailnum <- group_by(flights, tailnum)
by_tailnum2<- group_by(flights, year) # note difference from SQL! 

delay <- summarise(by_tailnum, count = n(),
                   dist = mean(distance, na.rm = TRUE),
                   delay = mean(arr_delay, na.rm = TRUE)) %>%
  filter(count > 20, dist < 2000)

ggplot(delay, aes(dist, delay)) + geom_point(aes(size = count), alpha = 1/2) +
  geom_smooth() +
  scale_size_area()
```

# Piping 101
```{r Piping}
flights %>%
  group_by(year, month, day) %>%
  select(arr_delay, dep_delay) %>%
  summarise(
    arr = mean(arr_delay, na.rm = TRUE),
    dep = mean(dep_delay, na.rm = TRUE)
  ) %>%
  filter(arr > 30 | dep > 30)
```

